Compiling refinement types to dependent types

### 2020-07-21 tex/paper.tex
More consistent \tau usage in core language
Lemma hyperrefing
Make the non-dependent pair labeled

\documentclass[a4paper]{article}
\usepackage{a4wide}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{ottalt}
\usepackage{subcaption}
\usepackage{cite,url,babelbib}
\usepackage[english]{babel}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

@@

Although the syntax doesn't mention $\tau$, we will use it as a metavariable
where a type (an inhabitant of $\star$) is expected.

@@

We also establish some derived forms in the core language to simplify the subsequent exposition,
and state some straightforward facts that will be useful later:
\begin{itemize}
  \item Simplified variable typing rule which we call \ranchor{CT-VarW}:
    \[
      \inferrule{\nonterm{GC} \vdash \nonterm{ec} : s \\ \gterm x : \nonterm{ec} \in \nonterm{GC}}
                {\nonterm{GC} \vdash \gterm x : \nonterm{ec}}.
    \]
    This rule unfolds into a single \ranchor{CT-Var} defining $x$,
    followed by a sequence of \ranchor{CT-Weaken}
    to ``move'' the $x$ into the right position in the final $\nonterm{GC}$.
  \item Non-dependent function type:
    \[
      \tau_1 \rightarrow \tau_2 \triangleq \Pi \_ : \tau_1. \tau_2.
    \]
  \item Dependent pair type, or $\Sigma$-type, with the first component being a value $x$ of type $\tau$ and the second component being of type $P x$:
    \[
      (x : \tau \dast P x) \triangleq \Pi \alpha : \star. (\Pi x : \tau. P x \rightarrow \alpha) \rightarrow \alpha,
    \]
    with the corresponding constructor (we overload the syntax here to resemble Idris):
    \[
      (x \dast p) \triangleq \lambda \alpha : \star. \lambda f : (\Pi x' : \tau. P x' \rightarrow \alpha). f x p
    \]
    and projections
    \[
      \text{fst} \triangleq \lambda \sigma : (x : \tau \dast P x). \sigma \tau (\lambda x' : \tau. \lambda \_ : P x'. x'),
    \]
    \[
      \text{snd} \triangleq \lambda \sigma : (x : \tau \dast P x). \sigma \tau (\lambda x' : \tau. \lambda p : P x'. p).
    \]
    It's easy to see from the formation rule that
    \begin{lemma}
      If $\nonterm{tc}$ and $P x$ are well-formed in $\nonterm{GC}$,
      then $(x : \nonterm{tc} \dast P x)$ is well-formed in $\nonterm{GC}$ and has type $\star$.
    \end{lemma}
  \item Non-dependent pair type:
    \[
      (\tau_1, \tau_2) \triangleq \{ x : \tau_1\ **\ (\lambda \_ : \tau_1. \tau_2) x \}.
    \]
    As a direct consequence of the previous lemma we obtain
    \begin{lemma}
      If $\nonterm{tc}_1, \nonterm{tc}_2$ are well-formed in $\nonterm{GC}$,
      then $(\nonterm{tc}_1, \nonterm{tc}_2)$ is well-formed in $\nonterm{GC}$ and has type $\star$.
    \end{lemma}
  \item Equality of expressions $\nonterm{ec}_1, \nonterm{ec}_2$ of type $S$:
    \[
      \nonterm{ec}_1 \equiv \nonterm{ec}_2 \triangleq \Pi P : S \rightarrow \star. (P x \rightarrow P y, P y \rightarrow P x).
    \]
    The reader might recognize this as the Leibniz equality \cite{FindSomethingForThis}.
  \item ADTs, expressed using an encoding similar to Boehm-Berarducci's \cite{Bohm85}.
    We borrow the syntax for the ADTs from the surface language and put a hat on it,
    so let's take some ADT $\widehat{\langle \overline{l_i : \nonterm{ts}_i}^i \rangle}$.
    The usual Boehm-Berarducci encoding associates the type
    $\Pi \alpha : \star. \overline{(\nonterm{ts}_i \rightarrow \alpha)}^i \rightarrow \alpha$ with it.

>>>>

\documentclass[a4paper]{article}
\usepackage{a4wide}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{hyperref,cleveref}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{ottalt}
\usepackage{subcaption}
\usepackage{cite,url,babelbib}
\usepackage[english]{babel}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\crefname{lemma}{Lemma}{Lemmas}

@@

Although the syntax doesn't mention $\nonterm{tc}$, we will use it as a metavariable
where a type (an inhabitant of $\star$) is expected.

@@

We also establish some derived forms in the core language to simplify the subsequent exposition,
and state some straightforward facts that will be useful later:
\begin{itemize}
  \item Simplified variable typing rule which we call \ranchor{CT-VarW}:
    \[
      \inferrule{\nonterm{GC} \vdash \nonterm{ec} : s \\ \gterm x : \nonterm{ec} \in \nonterm{GC}}
                {\nonterm{GC} \vdash \gterm x : \nonterm{ec}}.
    \]
    This rule unfolds into a single \ranchor{CT-Var} defining $x$,
    followed by a sequence of \ranchor{CT-Weaken}
    to ``move'' the $x$ into the right position in the final $\nonterm{GC}$.
  \item Non-dependent function type:
    \[
      \nonterm{tc}_1 \rightarrow \nonterm{tc}_2 \triangleq \Pi \_ : \nonterm{tc}_1. \nonterm{tc}_2.
    \]
  \item Dependent pair type, or $\Sigma$-type, with the first component being a value $x$ of type $\tau$ and the second component being of type $P x$:
    \[
      (x : \nonterm{tc} \dast P x) \triangleq \Pi \alpha : \star. (\Pi x : \nonterm{tc}. P x \rightarrow \alpha) \rightarrow \alpha,
    \]
    with the corresponding constructor (we overload the syntax here to resemble Idris):
    \[
      (x \dast p) \triangleq \lambda \alpha : \star. \lambda f : (\Pi x' : \nonterm{tc}. P x' \rightarrow \alpha). f x p
    \]
    and projections
    \[
      \text{fst} \triangleq \lambda \sigma : (x : \nonterm{tc} \dast P x). \sigma \nonterm{tc} (\lambda x' : \nonterm{tc}. \lambda \_ : P x'. x'),
    \]
    \[
      \text{snd} \triangleq \lambda \sigma : (x : \nonterm{tc} \dast P x). \sigma \nonterm{tc} (\lambda x' : \nonterm{tc}. \lambda p : P x'. p).
    \]
    It's easy to see from the formation rule that
    \begin{lemma}
      If $\nonterm{tc}$ and $P x$ are well-formed in $\nonterm{GC}$,
      then $(x : \nonterm{tc} \dast P x)$ is well-formed in $\nonterm{GC}$ and has type $\star$.
    \end{lemma}
  \item Non-dependent pair type:
    \[
      (\nonterm{tc}_1, \nonterm{tc}_2) \triangleq \{ x : \nonterm{tc}_1\ **\ (\lambda \_ : \nonterm{tc}_1. \nonterm{tc}_2) x \}.
    \]
    As a direct consequence of the previous lemma we obtain
    \begin{lemma}\label{lma:non_dep_pair_typing}
      If $\nonterm{tc}_1, \nonterm{tc}_2$ are well-formed in $\nonterm{GC}$,
      then $(\nonterm{tc}_1, \nonterm{tc}_2)$ is well-formed in $\nonterm{GC}$ and has type $\star$.
    \end{lemma}
  \item Equality of expressions $\nonterm{ec}_1, \nonterm{ec}_2$ of type $S$:
    \[
      \nonterm{ec}_1 \equiv \nonterm{ec}_2 \triangleq \Pi P : S \rightarrow \star. (P x \rightarrow P y, P y \rightarrow P x).
    \]
    The reader might recognize this as the Leibniz equality \cite{FindSomethingForThis}.
  \item ADTs, expressed using an encoding similar to Boehm-Berarducci's \cite{Bohm85}.
    We borrow the syntax for the ADTs from the surface language and put a hat on it,
    so let's take some ADT $\widehat{\langle \overline{l_i : \nonterm{ts}_i}^i \rangle}$.
    The usual Boehm-Berarducci encoding associates the type
    $\Pi \alpha : \star. \overline{(\nonterm{ts}_i \rightarrow \alpha)}^i \rightarrow \alpha$ with it.
